Towards practical dynamic induction control of wind farms: analysis of optimally controlled wind-farm boundary layers and sinusoidal induction control of first-row turbines

### 2018-02-14 draft.tex
Section --> Sect.

The current paper presents efforts on understanding optimal control dynamics observed in the optimal induction control simulations by \cite{munters2017optimal} (further denoted as MM17). The outline of the paper is as follows: first, Section~\ref{sec:optimal_control_sims} discusses the numerical setup and optimal control simulations of MM17 that will be further analysed in the current paper. Section~\ref{sec:analysis_controls_analysis} presents an analysis of the control and thrust force dynamics, and performs some numerical experiments to elucidate characteristics of the optimal controls. It will be shown that the coherent behavior of turbines situated in the first row of the wind farm stands out from their downstream counterparts. Hereafter, Section~\ref{sec:analysis_flow_vis} identifies the shedding of vortex rings from the first row based on a flow visualization. Further, a simple sinusoidal thrust control approach is presented that successfully mimics this process with a robust increase in power extraction extraction in the second row. Next, Section~\ref{sec:analysis_intermediate} shortly discusses the behavior of the intermediate rows, i.e. turbines which have both upstream and downstream neighbors, for which similar simple control strategies remain elusive thus far. In conclusion, Section~\ref{sec:analysis_summ} summarizes the main findings of this paper.

@@

First, the thrust coefficient signals themselves are analyzed in Section~\ref{sec:analysis_thrust}. Second, the optimized thrust coefficients are applied only to subsets of turbine rows in Section~\ref{sec:application_subset}. In this way, the interdependency of optimized thrust coefficients in different rows can be evaluated. Third, additional optimal control simulations, in which only one single active row is optimized, are discussed in Section~\ref{sec:optimization_single}. These optimizations provide an indication on how increased power potential is distributed among the rows, and allows to compare the resulting single-row optimized thrust coefficients with the fully cooperative coefficients from case MM17. Fourth, Section~\ref{sec:modified_thrust} evaluates the dependency of optimized thrust coefficients on the actual turbulent flow realization. Finally, Section~\ref{sec:analysis_discussion} discusses the main conclusions from the abovementioned sections, and summarizes the lessons learned.

@@

The current section further focuses on the analysis of the first row turbines. First, a qualitative analysis of the instantaneous flow field is performed in Section~\ref{sec:opt_visualization}, resulting in the observation of vortex rings being shed from first-row turbines. Thereafter, this mechanism is mimicked by imposing sinusoidally varying thrust coefficients in these turbines in Section~\ref{sec:opt_sinus}, with the aim of increasing power through similar mechanisms as in the computationally expensive optimal control cases.

>>>>

The current paper presents efforts on understanding optimal control dynamics observed in the optimal induction control simulations by \cite{munters2017optimal} (further denoted as MM17). The outline of the paper is as follows: first, Sect.~\ref{sec:optimal_control_sims} discusses the numerical setup and optimal control simulations of MM17 that will be further analysed in the current paper. Section~\ref{sec:analysis_controls_analysis} presents an analysis of the control and thrust force dynamics, and performs some numerical experiments to elucidate characteristics of the optimal controls. It will be shown that the coherent behavior of turbines situated in the first row of the wind farm stands out from their downstream counterparts. Hereafter, Sect.~\ref{sec:analysis_flow_vis} identifies the shedding of vortex rings from the first row based on a flow visualization. Further, a simple sinusoidal thrust control approach is presented that successfully mimics this process with a robust increase in power extraction extraction in the second row. Next, Sect.~\ref{sec:analysis_intermediate} shortly discusses the behavior of the intermediate rows, i.e. turbines which have both upstream and downstream neighbors, for which similar simple control strategies remain elusive thus far. In conclusion, Sect.~\ref{sec:analysis_summ} summarizes the main findings of this paper.

@@

First, the thrust coefficient signals themselves are analyzed in Sect.~\ref{sec:analysis_thrust}. Second, the optimized thrust coefficients are applied only to subsets of turbine rows in Sect.~\ref{sec:application_subset}. In this way, the interdependency of optimized thrust coefficients in different rows can be evaluated. Third, additional optimal control simulations, in which only one single active row is optimized, are discussed in Sect.~\ref{sec:optimization_single}. These optimizations provide an indication on how increased power potential is distributed among the rows, and allows to compare the resulting single-row optimized thrust coefficients with the fully cooperative coefficients from case MM17. Fourth, Sect.~\ref{sec:modified_thrust} evaluates the dependency of optimized thrust coefficients on the actual turbulent flow realization. Finally, Sect.~\ref{sec:analysis_discussion} discusses the main conclusions from the abovementioned sections, and summarizes the lessons learned.

@@

The current section further focuses on the analysis of the first row turbines. First, a qualitative analysis of the instantaneous flow field is performed in Sect.~\ref{sec:opt_visualization}, resulting in the observation of vortex rings being shed from first-row turbines. Thereafter, this mechanism is mimicked by imposing sinusoidally varying thrust coefficients in these turbines in Sect.~\ref{sec:opt_sinus}, with the aim of increasing power through similar mechanisms as in the computationally expensive optimal control cases.
